As-you-go" instead of "after-the-fact": A network approach to scholarly communication and evaluation

### 2018-03-30 docs/manuscript.Rmd
Update first use case

  We describe three use cases of network based evaluation to contextualize the ideas proposed above. For each use case, we first provide a general and non-exhaustive overview of the possibilities with network based evaluation. Subsequently, we specify a scenario for that use case, how an evaluation question flows from that scenario, how an indicator to answer that question could be operationalized, and how that indicator could inform the evaluation process. With these use cases we hope to illustrate that network based evaluation would align better with the implicit evaluation criteria already present in common research evaluation scenarios.

@@

  Funders of scholarly research have certain aims when distributing their financial resources amongst researchers. Currently, funders often use generic "one size fits all" indicators to evaluate the quality of researchers (e.g., JIF, h-index, citation counts). Given that funders often have specific aims with funding calls, these funding calls could form the basis of research evaluation. Potentially, given that network analysis allows different questions to be asked, funders might also change the aims of funding calls to shift towards more specific goals than generic notions of "innovation" or "discovery".

  One specific scenario is that of a funding agency calling for proposals to fund researchers to extend an already existing research line. This is not an implausible scenario, where funding agencies aim to fund several million dollars (or similar in other currencies) in order to increase follow through in research lines. A specific example might be the Dutch national funding agency "Vici" funding scheme, which aims to fund "senior researchers who have successfully demonstrated the ability to develop their own innovative lines of research" (https://perma.cc/GB83-RE4J).

  Whether researchers who submitted proposals actually have built a research line could be evaluated by looking at how interconnected each researcher's personal network is. Let us assume that a research line here would mean that new research efforts interconnect with previous efforts by that same researcher (i.e., building on previous work). Additionally, we could assume that building a research line means that the research line becomes more present in the network over the years. Building a research line thus could be reformulated into questions about the network of directly linked output and its development over time.

  Operationalizing the concept research line as increased interconnectedness of outputs over time, we could compute the network density of an applicant over the last five years to inform how the applicant's research aligns with the aim of the funding scheme. One way of computing density would be to tally the number of links and divide them by the number of possible links. By taking snapshots of the network of outputs of that researcher in the last five years on January 1st, we could compute an indicator to inform us about the development of the researcher's network of outputs.

  The development of network density over time could help inform the evaluation, but one measure could hardly be deemed the only decision criterion. As such, it only provides an indication as to whether an applicant aligns with the aim of the funding agency. Other questions still need to be answered by the committee. For example, is the project feasible or does the proposal extend the previous research line? Some of these other questions could also be seen as questions about the future development of the network and investigated, supplying a set of indicators with which to evaluate the applicant.

@@

  Universities use research evaluation for the internal allocation of resources and to hire new scientists. A research group within a university could use network analysis to assess how (dis)connected a group's outputs are or how their group compares to similar groups at other institutions.  Using network indicators, it would also become possible to assess whether a job applicant fulfills certain criteria, such as whether their outputs connect to existing outputs of a group. If a university wants to stimulate more diversity in research background, network analysis could also be used to identify those who are farther removed from the current researchers at the institution.  Considering that universities are often evaluated on the same generic indicators as individual researchers (e.g., JIF) in the rankings, such new and more precise evaluation tools might also help specify goals of a university and reduce the rat race. Like the use case for funders, network based evaluation allows universities to focus on question- and mission aligned evaluation.

@@

  Increased connectivity can be computed as a difference measure of the research group's network density with and without the applicant. In order to take into account the number of produced outputs, the computed density takes into account the number of outputs of an applicant. Moreover, the head stipulates that the minimum increase in network density needs to be five percentage points. To evaluate applicants, each gets a score that is made up of the difference between the current network density and the network density if they were hired. For example, baseline connectivity within a group might be 60%, hence, the network density has to be at least 65% for one of the applicants to pass the evaluation criterium.

>>>>

  We describe three use cases of network based evaluation to contextualize the ideas proposed above. For each use case, we first provide a general and non-exhaustive overview of the possibilities with network based evaluation. Subsequently, we specify a scenario for that use case, how an evaluation question flows from that scenario, how an indicator to answer that question could be operationalized, and how that indicator could inform the evaluation process. With these use cases we hope to illustrate that network based evaluation could align better with the implicit evaluation criteria already present in common research evaluation scenarios.

@@

  Funders of scholarly research often have specific aims when distributing their financial resources amongst researchers. Funders often use generic "one size fits all" indicators to evaluate the quality of researchers and research (e.g., JIF, h-index, citation counts). Given that funding calls often have specific aims, these funding calls could be used as the basis of research evaluation if we move beyond these generic measures. 
<!-- Potentially, given that network analysis allows different questions to be asked, funders might also change the aims of funding calls to shift towards more specific goals than generic notions of "innovation" or "discovery". -->

  A scenario could exist where a funding agency wants to fund researchers to extend an existing and interconnected research line. This is not an implausible scenario, where funding agencies aim to fund several million dollars (or similar in other currencies) in order to increase follow through in research lines. A specific example might be the Dutch national funding agency "Vici" funding scheme, which aims to fund "senior researchers who have successfully demonstrated the ability to develop their own innovative lines of research" (https://perma.cc/GB83-RE4J).

  Whether researchers who submitted proposals actually built a connected research line could be evaluated by looking at how interconnected each researcher's personal network of modules is. Let us assume that a research line here would mean that new research efforts interconnect with previous efforts by that same researcher (i.e., building on previous work). Additionally, we could assume that building a research line means that the research line becomes more present in the network over the years. Building a research line thus could be reformulated into questions about the network of directly linked output and its development over time.

  Operationalizing the concept 'research line' as increased interconnectedness of modules over time, we could compute the network density per year. One way of computing density would be to tally the number of links and divide them by the number of possible links. By taking snapshots of the network of outputs of that researcher in for example the last five years on January 1st, we could compute an indicator to inform us about the development of the researcher's network of outputs.

  The development of network density over time could help inform the evaluation, but one measure could hardly be deemed the only decision criterion. As such, it only provides an indication as to whether an applicant aligns with the aim of the funding agency. Other questions would still need to be answered by the evaluation committee. For example, is the project feasible or does the proposal extend the previous research line? Some of these other questions could also be seen as questions about the future development of the network and serve as their own questions to investigate the applicant on.

@@

  Universities can use research evaluation for the internal allocation of resources and to hire new scientists. As such, a research group within a university could apply network analysis to assess how (dis)connected a group's modules are or how their group compares to similar groups at other institutions. Using network indicators, it could become possible to assess whether a job applicant fulfills certain criteria, such as whether their modules connect to existing modules of a group. If a university wants to stimulate more diversity in research background, network analysis could also be used to identify those who are further removed from the current researchers at the institution. Considering that universities are often evaluated on the same generic indicators as individual researchers (e.g., JIF) in the rankings, such new and more precise evaluation tools might also help specify university goals. 
<!-- Like the use case for funders, network based evaluation allows universities to focus on question- and mission aligned evaluation.  -->

@@

  Increased connectivity can be computed as a difference measure of the research group's network density with and without the applicant. In order to take into account the number of produced modules, the computed density takes into account the number of modules of an applicant. Moreover, the head stipulates that the minimum increase in network density needs to be five percentage points. To evaluate applicants, each gets a score that is made up of the difference between the current network density and the network density if they were hired. For example, baseline connectivity within a group might be 60%, hence, the network density has to be at least 65% for one of the applicants to pass the evaluation criterium.
